---
title: "CART in R"
output: html_notebook
---

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

Set working directory, seed, load data.
```{r}
options(digits=2)
setwd("~/Desktop/MSA/Data Mining/Data")
load("breast_cancer.rdata")
set.seed(7515)
perm=sample(1:699)
BC_randomOrder=BCdata[perm,]
```

Split the data
```{r}
train = BC_randomOrder[1:floor(0.75*699),]
test = BC_randomOrder[(floor(0.75*699)+1):699,]
```

Next, build the default decision tree model using trophy as the metric.
`rpart` package is the most popular
```{r}
#install.packages("rpart")
library(rpart)
tree = rpart(Target ~ . - Target, data=train, method='class',
             parms = list(split='entropy'))
```

To display the decision tree, and label it, use the plot() and text() functions.
uniform=T causes all branches to have the sample length.
```{r}
.pardefault = par()
par(mai=c(.2, .2, .2, .2))
plot(tree, uniform=T)
text(tree)
#text(tree,use.n=T)
par(.pardefault)
```

Examine variable importance
```{r}
tree$variable.importance
```

Bar chart to visualize importance
```{r}
library('lattice')
barchart(tree$variable.importance[order(tree$variable.importance)],
         xlab = 'Importance', horiz=T, xlim=c(0,2000), ylab='Variable',
         main = 'Variable Importance', cex.names=0.8, las=2, col='orange')
```

R has `predict()` which is dope af
Output is a vector or matrix of predictions
We want to determine misclassification rate
```{r}
tscores = predict(tree, type='class')
scores = predict(tree, test, type='class')

cat('Training Misclassification Rate:', sum(tscores!=train$Target)/nrow(train))
```

Test Misclassification
```{r}
cat('Validation Misclassification Rate:', sum(scores!=test$Target)/nrow(train))
```

Visual Aesthetics
```{r}
#Fancy Tree Plot
#install.packages("rattle")
library(rattle)

#Enhanced Tree Plot
#install.packages("rpart")
library(rpart.plot)

# Color Selection for Fancy Plot
#install.packages("RColorBrewer")
library(RColorBrewer)

# Alternative Decision Tree Algorithm
#install.packages("party")
library(party)

# Convert rpart object to Binary Tree
#install.packages("partykit")
library(partykit)


```


Building graphs
```{r}
# Label Branches and Nodes with % of obs
prp(tree, type=3, extra=100)

#Label Branches and Nodes with Misclassification Rate
prp(tree, type=3, extra=2)

#Label Branches and Nodes with predicted probability of the class
prp(tree, type=3, extra=8)

# BEWARE: with Binary Trees where we typically want to show probability of success/failure for every node in the tree

prp(tree, type=0, extra=8, leaf.round=1, border.col=1,
    box.col=brewer.pal(10,"Set3")[tree$frame$yval], )
```


```{r}
# Lift Chart 
perf <- performance(pred, measure = "lift", x.measure = "rpp")
plot(perf, lwd = 3, colorize = TRUE, colorkey = TRUE,
     colorize.palette = rev(gray.colors(256)),
     main = "Lift Chart for Training Data")
abline(h = 1, lty = 3)
```

