---
title: "Sequence Analysis"
output: html_notebook
---

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

In Web Analytics: Called "Path Analytics"

Load the Data
```{r}
load("~/Desktop/MSA/Data Mining/Code/R/FaultCodes.rdata")
```

Use Seqecreate from TraMineR
* Undocumented
```{r}
#install.packages('TraMineR')
library(TraMineR)
#FCseqdata = seqecreate(id=FC$vehicleID, timestamp=FC$sequence, event=FC$Code)
#FCsubseqdata <- seqefsub(FCseqdata, pMinSupport=0.005)
#rules = TraMineR:::seqerules(FCsubseqdata)
head(rules)
```

Want to convert this to a probability/percentage
```{r}
#rules$Support = rules$Support/15075
rules=rules[order(-rules$Support),]
#rules=rules[order(-rules$Conf),]
#rules=rules[order(-rules$Lift),]
rules[1:10,]
```

In the top 10, we see a lot of A's and B's as the antecedents and various ones as subsequents

So, C's and F's were supposedly the worst. So we needed to look for things other than C and F in the antecedents so see if A's or B's could let us know that C's or F's were going to burn up before they actually do.