---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

# MLB Animated 'Race' Barchart
## Sam LaFell

Recently, I found an article by Keith McNulty (https://towardsdatascience.com/how-to-do-that-animated-race-bar-chart-57f3a8ff27a8) on Medium about making that animated race bar chat. Well, I followed along his guide and worked through it. Now, I want to make my own animated race barchart with my own data (actually, Lahman's Database Data).

So, I'm going to make the Race barchart that keeps track of which team has the most homeruns. This will be a cumulation of every players homeruns on that team for a given year. This will go from the 1800s to 2018.

This fits similarly with another Python project I did, https://samsdatajourney.wordpress.com/2019/02/16/my-first-data-project-from-data-retrieval-to-data-visualization/. Obviously this is in R and the end goal is different, but a similar idea.

We will have to do some cleaning and the visualization process will stay in R instead of exporting it to Tableau like I did with that one.

First, I download the data from http://www.seanlahman.com/baseball-archive/statistics/.

Next, we need to upload it. I'm working in RStudio Cloud at the moment. I also need to install all the packages and then call them.

```{r}
#install.packages("tidyverse")
#install.packages("knitr")
#install.packages("gganimate")
#install.packages("gifski")
#install.packages("png")
library(tidyverse)
library(knitr)
library(gganimate)
library(gifski)
library(png)
mlb <- read.csv("Batting.csv")
```

According to this History.com article, https://www.history.com/news/what-is-baseballs-modern-era, Baseballs Modern Era began in 1900. For this project, I am going to use 1900 as the first year and continue all the way to 2018. 

So, this means we need to create a new data.frame to only include years after 1900.
```{r filtering data to only include 1900 and later}
str(mlb)
mlb <- subset(mlb, mlb$yearID >= 1900)

# Next, we want to give us less variables to make this dataset easier to work with
mlb_less_columns <- select(mlb, c(playerID, yearID, teamID, lgID, HR))

str(mlb_less_columns)
```

So we are on the right track. Unfortunately, though, when we run `r str(mlb_less_columns)` we see that teamID has 149 levels? We know that there are only 30 teams in the MLB so 149 is definitely a bit much.

In doing my research on my last project, there are some team names that should be combined, and some that shouldn't. There are instances in which a Franchise moved cities, for example the Los Angeles Angels (current name) have been known as Anaheim Angels and a few other names.

Similarly, the Marlins were renamed, and the Expos moved to become the Washington Nationals. In the next chunk, I will rename the teams.

```{r reabbreviation the franchises to tally homerun totals appropriately}
mlb_new_teams <- mutate(mlb_less_columns, teamID = case_when(
  teamID == "SDN" | teamID == "SD" ~ "SD",
  teamID == "NYA" | teamID == "NYY" ~ "NYY",
  teamID == "NYN" | teamID == "NYM" ~ "NYM",
  teamID == "CHN" | teamID == "CHC" ~ "CHC",
  teamID == "SLN" | teamID == "STL" ~ "STL",
  teamID == "LAN" | teamID ==  "LAD" | teamID == "BRO" ~ "LAD",
  teamID == "SFN" | teamID == "SF" | teamID == "NYG" | teamID == "NY1" ~ "SF",
  teamID == "TBA" | teamID == "TB" | teamID == "TBR" | teamID == "TBD" ~ "TB",
  teamID == "ML4" | teamID == "MIL" ~ "MIL",
  teamID == "CAL" | teamID ==  "LAA" | teamID == "ANA" ~ "ANA",
  teamID == "FLO" | teamID ==  "MIA" ~ "MIA",
  teamID == "MON" | teamID ==  "WAS" ~ "WAS",
  teamID == "CIN" ~ "CIN",
  teamID == "BSN" | teamID == "MLN" | teamID == "ATL" ~ "ATL",
  teamID == "ARI" ~ "ARI",
  teamID == "SLB" | teamID == "MLA" | teamID == "BAL" ~ "BAL",
  teamID == "CHW" ~ "CHW",
  teamID == "CLE" ~ "CLE",
  teamID == "COL" ~ "COL",
  teamID == "DET" ~ "DET",
  teamID == "HOU" ~ "HOU",
  teamID == "KCR" | teamID == "KCA" ~ "KCR",
  teamID == "MIN" | teamID == "WSH" ~ "MIN",
  teamID == "PHA" | teamID == "OAK" ~ "OAK",
  teamID == "PHI" ~ "PHI",
  teamID == "PIT" ~ "PIT",
  teamID == "SEA" ~ "SEA",
  teamID == "WSA" | teamID == "TEX" ~ "TEX",
  teamID == "TOR" ~ "TOR",
))

str(mlb_new_teams)
```

With franchises moving, it's hard to know which teams belong together. Although, the baseball-reference page in my resources section really helped. I've dropped any teams that only existed for one year (most of those in 1914-1915 yearID) and only combined franchises with multiple years of existance.

Next, we need to drop any values included from those franchises that only existed for one year. They should have a null value for teamID right now.

```{r}
mlb_new_teams <- subset(mlb_new_teams, !is.na(teamID))

mlb_new_teams[2] <- lapply(mlb_new_teams[2], as.numeric)
mlb_new_teams[5] <- lapply(mlb_new_teams[5], as.numeric)
```


Now, our data is cleaned up and ready to go. Let's start manipulating.

## Manipulating the Data

Lucky, homeruns are calculated for each player for each year. All we need to do is group_by yearID and teamID and sum the homeruns.

```{r}
total_homeruns <- mlb_new_teams %>%
  group_by(teamID, yearID) %>%
  summarize(HR = sum(HR))
```

Now, for the chart to work like we want. We need to create rolling sums that update every year.

```{r creating rolling sums}
table <- total_homeruns %>% 
  dplyr::filter(yearID == 1900) %>% 
  dplyr::select(yearID, teamID, Homeruns = HR)
for (i in 1889:2017) {
  table <- total_homeruns %>% 
    dplyr::filter(yearID <= i) %>% 
    dplyr::group_by(teamID) %>% 
    dplyr::summarise(Homeruns = sum(HR, na.rm = TRUE)) %>% 
    dplyr::mutate(yearID = i) %>% 
    dplyr::bind_rows(table)
}
```

```{r interesting facts}
# add some historic facts to yearIDs
table <- table %>% 
  dplyr::mutate(
    year_label = dplyr::case_when(
      yearID <= 1900 ~ paste(yearID, "Beginning of the 'Modern Era' in MLB", sep = " - "),
      dplyr::between(yearID, 1914, 1918) ~ paste(yearID, "World War I", sep = " - "),
      dplyr::between(yearID, 1919, 1921) ~ paste(yearID, "Babe Ruth sold to the Yankees in 1920", sep = " - "),
      dplyr::between(yearID, 1926
                     , 1928) ~ paste(yearID, "Murderers' Row Yankees go 110-44 in 1927", sep = " - "),
      dplyr::between(yearID, 1930, 1933) ~ paste(yearID, "Babe Ruth's Called Shot in 1932", sep = " - "),
      dplyr::between(yearID, 1933, 1935) ~ paste(yearID, "First MLB All Star Game in 1933", sep = " - "),
      dplyr::between(yearID, 1940, 1943) ~ paste(yearID, "Joe DiMaggio Hits in 56 Straight Games in 1941", sep = " - "),
      dplyr::between(yearID, 1946, 1949) ~ paste(yearID, "Jackie Robinson Breaks the Color Barrier in 1947", sep = " - "),
      dplyr::between(yearID, 1950, 1952) ~ paste(yearID, "The Shot Heard Round the World in 1951", sep = " - "),
      dplyr::between(yearID, 1957, 1960) ~ paste(yearID, "Dodgers and Giants Move to California in 1958", sep = " - "),
      dplyr::between(yearID, 1960, 1962) ~ paste(yearID, "Mazeroski Wins WS Game 7 for Yankees on a Walk-Off Homerun in 1960", sep = " - "),
      dplyr::between(yearID, 1972, 1975) ~ paste(yearID, "Hank Aaron hits HR number 715", sep = " - "),
      dplyr::between(yearID, 1984, 1987) ~ paste(yearID, "Pete Rose Passes Ty Cobb in 1985 for Career Hits in 1985", sep = " - "),
      dplyr::between(yearID, 1994, 1997) ~ paste(yearID, "Cal Ripkin Passes Lou Gehrig for Consecutive Games in 1995", sep = " - "),
      dplyr::between(yearID, 2001, 2003) ~ paste(yearID, "Baseball Returns to New York after 9/11", sep = " - "),
      dplyr::between(yearID, 2004, 2006) ~ paste(yearID, "Red Sox win first World Series in 86 Years", sep = " - "),
      dplyr::between(yearID, 2006, 2008) ~ paste(yearID, "Barry Bonds with 756 Career Homeruns in 2007", sep = " - "),
      dplyr::between(yearID, 2008, 2009) ~ paste(yearID, "MLB Institutes Instant Reply on Home Runs in 2008", sep = " - "),
      dplyr::between(yearID, 2016, 2018) ~ paste(yearID, "The Cubs win their first World Series in 108 Years in 2016", sep = " - "),
      1L == 1L ~ as.character(yearID)
    )
  )
```

Now, Save it...

```{r}
save(table, file = "table.RData")
```


Now, the fun part!

# Creating the Animation

```{r retreiving top n ranking}
anim_table <- table %>%
  dplyr::group_by(yearID) %>%
  dplyr::mutate(
    rank = min_rank(-Homeruns) * 1,
    Value_rel = Homeruns / Homeruns[rank == 1],
    Value_lbl = paste0(" ", Homeruns)
  ) %>%
  dplyr::filter(rank <= 10) %>%
  dplyr::ungroup()
```

```{r}
p <- ggplot(anim_table, aes(rank)) +
  geom_tile(aes(y = Homeruns/2,
                height = Homeruns,
                width = 0.9,
                fill = "blue"
                ), alpha = 0.8, color = NA) +

  # text in x-axis (requires clip = "off" in coord_*)
  # paste(country, " ")  is a hack to make pretty spacing, since hjust > 1 
  #   leads to weird artifacts in text spacing.
  geom_text(aes(y = 0, label = paste(teamID, " ")), size = 12, vjust = 0.2, hjust = 1) +
  coord_flip(clip = "off", expand = FALSE) +
  scale_y_continuous(labels = scales::comma) +
  scale_x_reverse() +
  guides(color = FALSE, fill = FALSE) +

  labs(title='{closest_state}', x = "", y = "Total Homeruns") +
  theme(plot.title = element_text(hjust = 0, size = 30),
        axis.ticks.y = element_blank(),  # These relate to the axes post-flip
        axis.text.y  = element_blank(),  # These relate to the axes post-flip
        plot.margin = margin(2,2,1,16, "cm")) +

  transition_states(year_label, transition_length = 4, state_length = 1) +
  ease_aes('cubic-in-out')

animate(p, fps = 5, duration = 100, width = 2000, height = 1200, renderer = gifski_renderer("mlb_homeruns.gif"))
```

There you go!
There's room for the future to improve on this, i.e., spending more time verifying that the numbers are 100% accurate. I did some work on this for the project, but didn't want to spend too much time in nitty gritty details when my goal was to get the experience.

Thanks for reading, and let me know what you think!


Resources used:
https://stackoverflow.com/questions/27091292/change-column-value-by-row-string-value-in-r

https://www.baseball-reference.com/about/team_IDs.shtml

https://towardsdatascience.com/how-to-do-that-animated-race-bar-chart-57f3a8ff27a8

