---
title: "R Quiz 4"
output: html_notebook
---

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.


Import Libraries
```{r}
library(data.table)
library(dplyr)
library(stringr)
```


Load the Data and Change the CWD
```{r}
setwd("/Users/samlafell/Desktop/MSA/R/Class Data")
getwd()

df_auto <- fread("MPG_Data.csv", stringsAsFactors = FALSE)
```

Convert "Horsepower" to Numeric
```{r}
df_auto$Horsepower <- as.numeric(df_auto$Horsepower)
sum(is.na(df_auto$Horsepower))
```

#3 Using dplyr, add to "df_auto":
  # 3.1 Variable named "Car_Make" that is the first word from "Car_Name"
  # 3.2 Variable named "Car_Model" that is the remaining words (not first) from "Car_Name"
  # How many observations of "Car_Make" "ford" are there?
```{r}
df_auto <- df_auto %>%
  group_by(Car_Name) %>%
  mutate(Car_Make=str_split(Car_Name," ")[[1]][1]) %>% 
  mutate(Car_Model=paste(str_split(Car_Name," ")[[1]][-1], collapse=" ")) %>%
  ungroup(Car_Name)

sum(df_auto[10]=='ford')
```

#4 Using dplyr, how many cars are there for model year 73?
```{r}
df_auto %>%
  group_by(Model_Year, Car_Name) %>%
  filter(Model_Year==73)
```

#5 Using dplyr, what "Car_Make" "ford" car ("Car_Name") has the highest "Weight_per_Horsepower" (Weight/Horsepower) ratio? Ignore NAs.
```{r}
df_auto_ford <- df_auto %>%
  group_by(Car_Name) %>%
  mutate(Weight_per_Horsepower = mean(Weight/Horsepower, na.rm=TRUE)) %>%
  filter(Car_Make=='ford') %>%
  arrange(Weight_per_Horsepower) %>%
  ungroup(Car_Name)

df_auto
```


#6 Using dplyr, create a ranking variable, "MPG_Rank", that ranks, using the row_number() function, "MPG" by "Model_Year" in descending order.
  # What is the third highest "MPG" "Car_Name" for "Model_Year" 1972?
```{r}
df_auto_new3<-df_auto %>%
 arrange(Model_Year, desc(MPG))%>%
 group_by(Model_Year)%>%
 mutate(MPG_Rank=row_number())%>%
 ungroup(Model_Year)

filter(df_auto_new3, Model_Year=='72')
```


#7 Using dplyr and stringr, update "df_auto" so all values of "Car_Make" that contain "chev" say "chevrolet".
  #How many observations have "Car_Make" "chevrolet"?
```{r}
current <- 'chev'
new <- 'chevrolet'

df_auto <- df_auto %>%
  mutate(Car_Make = ifelse(str_detect(Car_Make, current), new, Car_Make))

print(length(which(df_auto$Car_Make=='chevrolet')))
```

#8 What is the average MPG for "chevrolet" cars?
```{r}
df_auto %>%
  group_by(Car_Make) %>%
  summarise(AvgMPG = mean(MPG))
```


#9 Using dplyr joins, what is the average "MPG" for car makes where "First_Year" exists?
```{r}
Car_Make <- c("amc", "audi", "bmw", "buick", "chevrolet", "datsun", "dodge", "ford")
First_Year <- c(1954, 1910, 1916, 1903, 1911, 1931, 1900, 1903)
df_auto_start <- data.frame(Car_Make=Car_Make, First_Year=First_Year, stringsAsFactors = FALSE)

df_auto_hold <- df_auto %>%
  right_join(df_auto_start, by="Car_Make")

mean(df_auto_hold$MPG)
```

#10 Using dplyr joins, how many cars exist in the data set where "First_Year" is null?
```{r}
# Left_join 
df_auto_hold <- df_auto %>%
  left_join(df_auto_start, by="Car_Make")

# note, we've retained all of the records but it's placed "NA" in the column where it couldn't find a matching record
dim(df_auto) 
dim(df_auto_hold) 

sum(is.na(df_auto_hold$First_Year))
```



