---
title: "R Notebook"
output: html_notebook
---

```{r}
plot(cars)
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

1 Prepare by:
  #1.1 Import the required libraries:
  #1.2 Set your working directory and upload the "Auto_MPG.csv" data. Ensure stringAsFactors=FALSE.
  #1.3 Create a column names "Car_Make" from the first element of the "Car_Name" variable
  #1.4 Update "horsepower" to be type numeric
  #1.4 How many unique values of "Car_Make" are there?
  
```{r}
library(data.table)
library(dplyr)
library(ggplot2)
library(stringr)

setwd("/Users/samlafell/Desktop/MSA/R/Class Data/")

df_auto <- fread("MPG_Data.csv", stringsAsFactors = FALSE)

df_auto <- df_auto %>%
  group_by(Car_Name) %>%
  mutate(Car_Make=str_split(Car_Name," ")[[1]][1]) %>% 
  mutate(Car_Model=paste(str_split(Car_Name," ")[[1]][-1], collapse=" ")) %>%
  ungroup(Car_Name)

df_auto$Horsepower <- as.numeric(df_auto$Horsepower)

n_distinct(df_auto$Car_Make)


```


We are going to create a function that provides am univariate analysis on a provided variable vector
    
#2 Change the type of "Cylinders", "Model_Year", and "Origin" from numeric to character.
    # How many variables in "df_auto" are of type character?

```{r}
df_auto$Cylinders <- as.character(df_auto$Cylinders)
df_auto$Model_Year <- as.character(df_auto$Model_Year)
df_auto$Orgin <- as.character(df_auto$Orgin)

str(df_auto)

sum(is.character(df_auto))
class(df_auto)
typeof(df_auto$MPG)
names(df_auto)
typeof(df_auto$MPG)
typeof(df_auto$Cylinders)
typeof(df_auto$Displacement)
typeof(df_auto$Horsepower)
typeof(df_auto$Weight)
typeof(df_auto$Acceleration)
typeof(df_auto$Model_Year)
typeof(df_auto$Orgin)
typeof(df_auto$Car_Name)
typeof(df_auto$Car_Make)
typeof(df_auto$Car_Model)
```


#3 Create a function "univariate_analysis" that:
    #3.1 Calculates the number of observations, mean, standard deviation, and quartiles of a provided vector. 
      #3.1 Ensure the na.rm argument is TRUE
    #3.2 Have the function return the results in a list with named elements "obs", "mean", "stdev", and "quartiles"
    #Using the function created, what is the 2nd quartile of "MPG"?


# 6 Thus far, we've assumed the variable provided will be numeric. However, we should account for categorical variables. 
  # Update the "univariate_analysis" function to:
    #6.1 If the provided vector is numeric, execute everything created in the function thus far
    #6.2 If the provided vector is character:
      #6.2.1 Store the number of elements in the vector in a variable named "obs"
      #6.2.2 Store the unique values of the vector in a variable named "unique_levels" 
      #6.2.1 Store a bar chart in a variable "bar_chart" that plots the number of observations by the unique levels of the variable 
    #6.3 Returns the previous variables in a list with the elements named "obs", "unique_levels", and "bar_chart")
    #Using the function created, what "Model_Year" has the greatest number of observations?
```{r}
univariate_analysis <- function(data){
  
  if(is.numeric(data)==FALSE){
     obs <- length(data)
     unique_levels <- n_distinct(data)
     data <- as.data.frame(data)
     bar_chart <- ggplot(data, aes(x=data)) +
       geom_bar()
     
     l2 <- list(n_obs=obs, unique_levels=unique_levels, bar_chart=bar_chart)
     return(l2)
  }
  
  n <- length(data)
  mean <- mean(data, na.rm = TRUE)
  std <- sd(data, na.rm = TRUE)
  percentiles  <- quantile(data, probs=c(0, 0.25, 0.50, 0.75, 1), na.rm = TRUE)
  data <- as.data.frame(data)
  
  histogram <- ggplot(data, aes(x=data)) +
    geom_histogram(bins=20) +
    geom_vline(aes(xintercept = quantile(data)[2]), linetype = "dashed", colour = "red") +
    geom_vline(aes(xintercept = quantile(data)[3]), linetype = "dashed", colour = "blue") +
    geom_vline(aes(xintercept = quantile(data)[4]), linetype = "dashed", colour = "green") +
    geom_text(aes(x=quantile(data)[2]-1.2, label="25% quartile", y=20), colour="red", angle=90, text=element_text(size=12)) +
    geom_text(aes(x=quantile(data)[3]-1.2, label="50% quartile", y=32), colour="blue", angle=90, text=element_text(size=12)) +
    geom_text(aes(x=quantile(data)[4]-1.2, label="75% quartile", y=20), colour="green", angle=90, text=element_text(size=12)) +
    theme_minimal()
  
  
  l1 <- list(n_obs=n, mean=mean, std=std, percentiles=percentiles, histogram=histogram)
  return(l1)
  
}

univariate_analysis(df_auto$Model_Year)

```


#4 To the "univariate_analysis" function, create a histogram plot named "histogram". Add the plot to the return list with the element named "histogram". 
  #Note, you will be providing your function a vector, but ggplot only accepts data frames. You will need to convert your vector to a data frame with one column.
  #What is the default number of bins in the histogram for "MPG"?

```{r}
univariate_analysis(df_auto$MPG)
```

 #6.3 Returns the previous variables in a list with the elements named "obs", "unique_levels", and "bar_chart")
    #Using the function created, what "Model_Year" has the greatest number of observations?
```{r}

univariate_analysis(df_auto$Model_Year)


```



#7 Using the "univariate_analysis" function, what is the standard deviation of "Horsepower" 
```{r}
univariate_analysis(df_auto$Horsepower)

sd(df_auto$Horsepower, na.rm = TRUE)

sapply(df_auto, sd, na.rm = TRUE)

```


#8 Using the "univariate_analysis" function, what is the median of "Displacement" 
```{r}
univariate_analysis(df_auto$Displacement)

```



#9 Using the "univariate_analysis" function, what is the most common number of cylinders for cars in the data set?
```{r}
univariate_analysis(df_auto$Cylinders)

```




#10 Using the "univariate_analysis" function, what is the mean of Weight? 
```{r}
univariate_analysis(df_auto$Weight)
```