---
title: "R Quiz 6"
output: html_notebook
---

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.


#1 Load the following libraries:

  #1.1 Assign the "mpg" dataset from the ggplot2 library to a new data frame named "df_mpg"

  # How many observations are there in "df_mpg"

```{r}
library(ggplot2)
df_mpg <- data.frame(ggplot2::mpg)
dim(df_mpg)
```


#2 Use the cor function to determine correlations between the numeric variables in "df_mpg." What is the highest correlation, rounded to 2 digits?
```{r}
cor_matrix <- cor(select_if(df_mpg, is.numeric))
View(cor_matrix)
max(cor_matrix)
```

#3 Remove the variables "manufacturer", "model", "hwy", and "fl" from "df_mpg". How many columns are in "df_mpg"?
```{r}
df_mpg2 <- df_mpg[-c(1, 2, 9, 10)]
colnames(df_mpg2)
dim(df_mpg2)
```

#4 Test the hypothesis that cars built in "year" "1999" have no difference in "cty" mpg as cars built in "year" "2008" using a Welch's two sample t test.

  #What is the p-value of the t test, rounded to two digits?

```{r}
t.result <- t.test(cty ~ year, data=filter(mpg, year %in% c('1999','2008')))
t.result
```

#5 Spit the data into 70% training, "df_mpg_t", and 30% validation, "df_mpg_v" data sets.

  #5.1 Use set.seed(1)

  #5.2 Use createDataPartition from Caret, splitting on the variable "cty"

  # How many observations are in "df_mpg_t"? 
```{r}
set.seed(1)
train_ind <- createDataPartition(mpg$cty, p=0.70, list=FALSE)
df_mpg_t <- df_mpg2[train_ind,]
df_mpg_v <- df_mpg2[-train_ind,]
dim(df_mpg_t)
```

#6 Create a linear regression model with response variable "cty" and independent variable "displ" from "df_mpg_t". 

  #6.1 assign the model output to a variable "m1_fit"

  #What is the Adj R2 of the model, rounded to 2 digits?
```{r}
m1_fit <- lm(cty~displ, 
               data=df_mpg_t, 
               method="lm")
summary(m1_fit)
```

#7 Test the hypothesis that the residuals from "m1_fit" are normally distributed using a Shapiro test. 

  #Is the null hypothesis rejected?
  #Shapiro test - remember null hypothesis is distribution is normal

```{r}
shapiro.test(m1_fit$residuals)
```


#8 Fit a model, "m2_fit" that is a linear regression model with response variable "cty" and independent variables "displ" and "class".

  #Using Adj R2 as a evaluation criteria, which model, "m1_fit" or "m2_fit" fits the data better?
```{r}
m2_fit <- lm(cty~displ+class, 
               data=df_mpg_t, 
               method="lm")

summary(m2_fit)
```


#9 Score the validation data "df_mpg_v" with both models "m1_fit" and "m2_fit". 

  #Which model, "m1_fit" or "m2_fit" performed better using RMSE as a evaluation criteria? Use the postResample function from Caret to determine the RMSE.
```{r}
library(caret)
#Use the predict function 
scored <- predict(m1_fit, df_mpg_v)
scored

#Evaluate the scoring
postResample(scored, df_mpg_v$cty)


#Use the predict function
scored <- predict(m2_fit, df_mpg_v)
scored

#Evaluate the scoring
postResample(scored, df_mpg_v$cty)
```
