---
title: "R Notebook"
output: html_notebook
---


```{r}
## Cancer Genetics DNA

load('~/Desktop/MSA/Linear Algebra/Data/geneCancerUCI.rdata')
table(cancerlabels$Class)
```

Original Source: The cancer genome atlas pan-cancer analysis project

BRCA = Breast Invasive Carcinoma
COAD = Colon Adenocarcinoma
KIRC = Kidney Renal clear cell Carcinoma
LUAD = Lung Adenocarcinoma
PRAD = Prostate Adenocarcinoma
We are going to want to plot the data points according to their different classification labels. We should pick out a nice color palette for categorical attributes.

```{r}
library(RColorBrewer)
display.brewer.all()
```


```{r}
palette(brewer.pal(n = 8, name = "Dark2"))
```

The first step is typically to explore the data. Obviously we can’t look at ALL the scatter plots of input variables. For the fun of it, let’s look at a few of these scatter plots which we’ll pick at random. First pick two column numbers at random, then draw the plot, coloring by the label.


```{r}
randomColumns = sample(2:20532,2)
plot(cancer[,randomColumns],col = cancerlabels$Class)
```

The prcomp function is the one I most often recommend for reasonably sized principal component calculations in R. This function returns a list with class “prcomp” containing the following components (from help prcomp):

sdev: the standard deviations of the principal components (i.e., the square roots of the eigenvalues of the covariance/correlation matrix, though the calculation is actually done with the singular values of the data matrix).
rotation: the matrix of variable loadings (i.e., a matrix whose columns contain the eigenvectors). The function princomp returns this in the element loadings.
x: if retx is true the value of the rotated data (the centred (and scaled if requested) data multiplied by the rotation matrix) is returned. Hence, cov(x) is the diagonal matrix diag(sdev^2). For the formula method, napredict() is applied to handle the treatment of values omitted by the na.action.
center, scale: the centering and scaling used, or FALSE.
Now let’s compute the first two principal components and examine the data projected onto these axes. We can then look in 3 dimensions.

```{r}

pcaOut = prcomp(cancer[,2:20532],3, scale = F)
plot(pcaOut$x[,1],pcaOut$x[,2],col = cancerlabels$Class, xlab = "Principal Component 1", ylab = "Principal Component 2", main = 'Genetic Samples Projected into 2-dimensions')
```

3-dimensional plot. Problem is to get the plot points colored by group, we need to write a function that creates a vector of colors

```{r}
colors = factor(palette())
colors = colors[cancerlabels$Class]
```

Need rgl library
```{r}
#install.packages('rgl')
library(rgl)
plot3d(x = pcaOut$x[,1], y = pcaOut$x[,2],z= pcaOut$x[,3],col = colors, xlab = "Principal Component 1", ylab = "Principal Component 2", zlab = "Principal Component 3")
```


Proportion of Variance explained by 2 components:

```{r}
sum(pcaOut$sdev[1:2]^2)/sum(pcaOut$sdev^2)
```

Proportion of Variance explained by 3 components:

```{r}
sum(pcaOut$sdev[1:3]^2)/sum(pcaOut$sdev^2)
```
